---
title: "march"
author: "Josh LaFronz"
date: "2024-03-17"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(scales)
library(ggthemes)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(lubridate)
library(stringr)

# model building
library(MASS)
library(leaps)
library(car)
library(olsrr)
```

## Tournament Field

East:

1 UCONN - 16 Stetson

8 FAU - 9 Northwestern

5 SDSU - 12 UAB

4 Auburn - 13 Yale

6 BYU - 11 Duquesne

3 Illinois - 14 Morehead St

7 Washington St - 10 Drake

2 Iowa St - 15 South Dakota St

South:

1 Houston - 16 Longwood

8 Nebraska - 9 Texas A&M

5 Wisconsin - 12 James Madison

4 Duke - 13 Vermont

6 Texas Tech - 11 NC State

3 Kentucky - 14 Oakland

7 Florida - 10 Boise St/Colorado (FIRST FOUR)

2 Marquette - 15 Western Kentucky

Midwest:

1 Purdue - 16 Montana St/Grambling (FIRST FOUR)

8 Utah St - 9 TCU

5 Gonzaga - 12 McNeese St

4 Kansas - 13 Samford

6 South Carolina - 11 Oregon

3 Creighton - 14 Akron

7 Texas - 10 Virginia/Colorado St (FIRST FOUR)

2 Tennessee - 15 Saint Peter’s

West:

1 UNC - 16 Howard/Wagner (FIRST FOUR)

8 Mississippi St - 9 Michigan St

5 SMC - 12 GCU

4 Alabama - 13 Charleston

6 Clemson - 11 New Mexico

3 Baylor - 14 Colgate

7 Dayton - 10 Nevada

2 Arizona - 15 Long Beach St


## Notes gathered so far

Virginia is an abysmal free throw shooting team (First Four against Colorado)

Texas A&M is the worst three point shooting team (Against Nebraska)

```{r}
tourney_teams = c('Connecticut', 'Houston', 'Purdue',
'North Carolina', 'Iowa State', 'Marquette',
'Tennessee','Arizona','Illinois','Kentucky',
'Creighton','Baylor','Auburn','Duke','Kansas',
'Alabama','San Diego State','Wisconsin','Gonzaga',
"Saint Mary's (CA)", 'Brigham Young', 'Texas Tech',
'South Carolina','Clemson','Washington State', 'Florida','Texas','Dayton', 'Florida Atlantic',
'Nebraska','Utah State','Mississippi State',
'Northwestern','Texas A&M','TCU','Michigan State',
'Drake','Nevada','Duquesne','NC State', 'Oregon',
'New Mexico','UAB','James Madison','McNeese State',
'Grand Canyon','Yale','Vermont','Samford','College of Charleston','Morehead State','Oakland','Akron',
'Colgate','South Dakota State','Western Kentucky',
"Saint Peter's","Long Beach State","Stetson",
"Longwood")

first_four = c("Boise State", "Colorado", "Montana State", "Grambling",
               "Virginia","Colorado State",
               "Howard","Wagner")

conference = c("Big East", "Big 12", "Big Ten","ACC","Big 12","Big East","SEC",
               "Pac-12","Big Ten", "SEC", "Big East", "Big 12","SEC", "ACC","Big 12",
               "SEC","MWC","Big Ten","WCC","WCC","Big 12","Big 12","SEC",
               "ACC","Pac-12","SEC","Big 12","A-10","AAC","Big Ten","MWC","SEC",
               "Big Ten","SEC","Big 12","Big Ten","MVC","MWC","A-10","ACC","Pac-12",
               "MWC","AAC","Sun Belt","Southland","WAC","Ivy","AEC","Southern","CAA",
               "OVC","Horizon","MAC","Patriot","Summit","C-USA","MAAC","Big West","A-Sun",
               "Big South")

ff_conf = c("MWC","Pac-12","Big Sky","SWAC","ACC","MWC","MEAC","NEC")
```

```{r}
tourney_teams |>
  sort()
```

```{r}
seeding = data.frame(
  seed = c(sort(rep(seq(1,9),4)), 10,10, sort(rep(seq(11,15),4)), 16, 16),
  region = c(rep(c("East","South","Midwest","West"),9), "East", "West", rep(c("East","South","Midwest","West"),5), "East", "South"),
  team = tourney_teams,
  conference = conference,
  first_four = FALSE
)
seeding |> rbind(
data.frame(
  seed = c(10,10,16,16,10,10,16,16),
  region = c("South", "South", rep("Midwest", 4), "West", "West"),
  team = first_four,
  conference = ff_conf,
  first_four = TRUE
)) -> field_of_68
field_of_68
rm(conference)
rm(ff_conf)
rm(first_four)
rm(tourney_teams)
rm(seeding)
```
```{r}
# remove schools that lost in the first four
field_of_68 |> filter(team %in% c("Boise State","Montana State","Virginia","Howard") == FALSE) -> field_of_68
```

```{r}
# testing = read_csv("/Users/joshlafronz/Documents/my stuff/projects/datasets/march/spreftest.csv")
# glimpse(testing)
```


```{r}
# names(testing)
# testing |>
#   rename(tot_W = W...4, tot_L = L...5, conf_W = W...10, conf_L = L...11, home_W
#          = W...13, home_L = L...14, away_W = W...16, away_L = L...17,
#          PS = Tm., PA = Opp., team = School) |>
#   subset(select = -c(Rk, ...9, ...12, ...15, ...18, ...21)) |>
#   # filter(School %in% tourney_teams | School %in% tourney_teams)
#   merge(field_of_68, by = "team")
```

```{r}
field_of_68 |> group_by(conference) |> count()
```

```{r}
basic_off = read_csv("/Users/joshlafronz/Documents/my stuff/projects/datasets/march/basic_off.csv")
basic_def = read_csv("/Users/joshlafronz/Documents/my stuff/projects/datasets/march/basic_def.csv")

basic_off |>
  rename(tot_W = W...4, tot_L = L...5, conf_W = W...10, conf_L = L...11, home_W
         = W...13, home_L = L...14, away_W = W...16, away_L = L...17,
         PS = Tm., PA = Opp., team = School) |>
  subset(select = -c(Rk, ...9, ...12, ...15, ...18, ...21)) |>
  # filter(School %in% tourney_teams | School %in% tourney_teams)
  merge(field_of_68, by = "team") |> relocate(seed, .before = team) |>
  relocate(region, .before = team) |> relocate(conference, .before = G) -> basic_off

basic_def |>
  rename(tot_W = W...4, tot_L = L...5, conf_W = W...10, conf_L = L...11, home_W
         = W...13, home_L = L...14, away_W = W...16, away_L = L...17,
         PS = Tm., PA = Opp., team = School) |>
  subset(select = -c(Rk, ...9, ...12, ...15, ...18, ...21)) |>
  # filter(School %in% tourney_teams | School %in% tourney_teams)
  merge(field_of_68, by = "team") |> relocate(seed, .before = team) |> 
  relocate(seed, .before = team) |> relocate(conference, .before = G) -> basic_def
```

```{r}
basic_off |> mutate(
  X3pr = X3PA/FGA
) |> ggplot(aes(x = X3pr, y = X3P_pct)) + geom_boxplot()

basic_off |> mutate(
  X3pr = X3PA/FGA
) |> filter(X3P_pct > .4 | X3P_pct < .3)
```

```{r}
basic_off |> mutate(
  FTR = FTA/FGA
) |> ggplot(aes(x = FTR, y = FT_pct)) + geom_point()

basic_off |> mutate(
  FTR = FTA/FGA
) |> filter(FT_pct < .64)

basic_off |> mutate(
  FTR = FTA/FGA
) |> filter(FTR > .4)

basic_off |> mutate(
  FTR = FTA/FGA
) |> filter(FT_pct > .76)
```
```{r}
basic_off |> filter(tot_W > 29)
```



```{r}
advanced_off = read_csv("/Users/joshlafronz/Documents/my stuff/projects/datasets/march/advanced_off.csv")
advanced_def = read_csv("/Users/joshlafronz/Documents/my stuff/projects/datasets/march/advanced_def.csv")

advanced_off |>
  rename(tot_W = W...4, tot_L = L...5, conf_W = W...10, conf_L = L...11, home_W
         = W...13, home_L = L...14, away_W = W...16, away_L = L...17,
         PS = Tm., PA = Opp., team = School) |>
  subset(select = -c(Rk, ...9, ...12, ...15, ...18, ...21)) |>
  # filter(School %in% tourney_teams | School %in% tourney_teams)
  merge(field_of_68, by = "team") |> relocate(seed, .before = team) |>
  relocate(seed, .before = team) |> relocate(conference, .before = G) -> advanced_off

advanced_def |>
  rename(tot_W = W...4, tot_L = L...5, conf_W = W...10, conf_L = L...11, home_W
         = W...13, home_L = L...14, away_W = W...16, away_L = L...17,
         PS = Tm., PA = Opp., team = School, DRtg = ORtg) |>
  subset(select = -c(Rk, ...9, ...12, ...15, ...18, ...21)) |>
  # filter(School %in% tourney_teams | School %in% tourney_teams)
  merge(field_of_68, by = "team") |> relocate(seed, .before = team) |>
  relocate(seed, .before = team) |> relocate(conference, .before = G) -> advanced_def
```

```{r}
advanced_def |> filter(DRtg < 95)
```

```{r}
advanced_off |> left_join(advanced_def, by = (c('seed', 'team', 'conference'))) |> subset(select=c(seed, team, ORtg, DRtg))  |> mutate(
  net_rtg = ORtg - DRtg
) |> arrange(desc(net_rtg))
```

```{r}
advanced_off |> left_join(advanced_def, by = (c('seed', 'team', 'conference','region','first_four'))) |>
  subset(select=c(seed, team, region, ORtg, eFG_pct.x, TOV_pct.x, ORB_pct.x, FT_FGA.x, DRtg, eFG_pct.y,
                  TOV_pct.y, ORB_pct.y, FT_FGA.y)) |>
  # mutate(
  #   seed = fct_relevel(as.factor(seed), "1","16","8","9","5","12","4","13","6","11","3","14","7","10","2","15")
  # )
  mutate(
    seed = factor(seed, levels = c("1","16","8","9","5","12","4","13","6","11","3","14","7","10","2","15")),
    region = factor(region, levels = c("East", "South", "Midwest", "West"))
  ) -> four_factors
four_factors[order(four_factors$region, four_factors$seed),] -> four_factors
four_factors
```
```{r}
four_factors |> mutate(
  matchup = case_when(
    (seed == 1 | seed == 16) ~ "1-16",
    (seed == 2 | seed == 15) ~ "2-15",
    (seed == 3 | seed == 14) ~ "3-14",
    (seed == 4 | seed == 13) ~ "4-13",
    (seed == 5 | seed == 12) ~ "5-12",
    (seed == 6 | seed == 11) ~ "6-11",
    (seed == 7 | seed == 10) ~ "7-10",
    (seed == 8 | seed == 9) ~ "8-9",
  )
) |> relocate(matchup, .before = ORtg) -> four_factors
ggplot(four_factors, aes(x = ORtg, y = eFG_pct.x, color = matchup, group = matchup)) + geom_point(size = 4) +
  facet_wrap(~region) + geom_line()
```
```{r}
advanced_off |> left_join(advanced_def, by = (c('seed', 'team', 'conference','region','first_four'))) |>
  subset(select=c(seed, team, region, Pace.x, ORtg, eFG_pct.x, TOV_pct.x, ORB_pct.x, FT_FGA.x, DRtg, eFG_pct.y,
                  TOV_pct.y, ORB_pct.y, FT_FGA.y)) |> mutate(
  higher_seed = case_when(
    seed < 9 ~ TRUE,
    TRUE ~ FALSE
  ),
  matchup = case_when(
    (seed == 1 | seed == 16) ~ "1-16",
    (seed == 2 | seed == 15) ~ "2-15",
    (seed == 3 | seed == 14) ~ "3-14",
    (seed == 4 | seed == 13) ~ "4-13",
    (seed == 5 | seed == 12) ~ "5-12",
    (seed == 6 | seed == 11) ~ "6-11",
    (seed == 7 | seed == 10) ~ "7-10",
    (seed == 8 | seed == 9) ~ "8-9",
  )
) |> relocate(matchup, .before = ORtg) -> four_factors

four_factors

ggplot(four_factors, aes(x = eFG_pct.x, y = FT_FGA.x, color = matchup, group = matchup)) + geom_point(aes(size = ORtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()

ggplot(four_factors, aes(x = TOV_pct.x, y = ORB_pct.x, color = matchup, group = matchup)) + geom_point(aes(size = ORtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()

ggplot(four_factors, aes(x = eFG_pct.y, y = FT_FGA.y, color = matchup, group = matchup)) + geom_point(aes(size = DRtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()

ggplot(four_factors, aes(x = TOV_pct.y, y = ORB_pct.y, color = matchup, group = matchup)) + geom_point(aes(size = DRtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()
```
```{r}
advanced_off |> mutate(
  higher_seed = case_when(
    seed < 9 ~ TRUE,
    TRUE ~ FALSE
  ),
  matchup = case_when(
    (seed == 1 | seed == 16) ~ "1-16",
    (seed == 2 | seed == 15) ~ "2-15",
    (seed == 3 | seed == 14) ~ "3-14",
    (seed == 4 | seed == 13) ~ "4-13",
    (seed == 5 | seed == 12) ~ "5-12",
    (seed == 6 | seed == 11) ~ "6-11",
    (seed == 7 | seed == 10) ~ "7-10",
    (seed == 8 | seed == 9) ~ "8-9",
  )
) |> relocate(matchup, .before = ORtg) |> group_by(region, matchup, team) |> reframe(Pace = Pace)
```

```{r}
four_factors |> ggplot(aes(x = ORtg, y = eFG_pct.x, color = seed)) + geom_point(size = 4) +
  facet_wrap(~region)
```

```{r}
r1_winners = c("Connecticut","Florida Atlantic", "San Diego State","Auburn","Houston",
               "Texas A&M","James Madison","Duke","Brigham Young","Illinois","Drake",
               "Iowa State","Texas Tech", "Kentucky", "Colorado", "Western Kentucky",
               "North Carolina","Mississippi State","Saint Mary's (CA)","College of Charleston",
               "Purdue","Utah State","Gonzaga","Samford","New Mexico","Baylor","Nevada",
               "Arizona","Oregon","Creighton","Colorado State","Tennessee")

four_factors |> filter(team %in% r1_winners) -> four_factors
four_factors |>
  mutate(
    matchup = case_when(
      seed %in% c(1,16,8,9) ~ "1/16 vs 8/9",
      seed %in% c(5,12,4,13) ~ "5/12 vs 4/13",
      seed %in% c(6,11,3,14) ~ "6/11 vs 3/14",
      seed %in% c(7,10,2,15) ~ "7/10 vs 2/15"
    )
  ) -> four_factors

four_factors

ggplot(four_factors, aes(x = eFG_pct.x, y = FT_FGA.x, color = as.factor(seed), group = matchup)) + geom_point(aes(size = ORtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()

ggplot(four_factors, aes(x = TOV_pct.x, y = ORB_pct.x, color = as.factor(seed), group = matchup)) + geom_point(aes(size = ORtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()

ggplot(four_factors, aes(x = eFG_pct.y, y = FT_FGA.y, color = as.factor(seed), group = matchup)) + geom_point(aes(size = DRtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()

ggplot(four_factors, aes(x = TOV_pct.y, y = ORB_pct.y, color = as.factor(seed), group = matchup)) + geom_point(aes(size = DRtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()
```
```{r}
r2_winners = c("Connecticut","Auburn","Illinois","Iowa State","Houston","James Madison","Kentucky",
               "Colorado","North Carolina","Saint Mary's (CA)","New Mexico","Arizona","Purdue",
               "Samford","Creighton","Tennessee")

four_factors |> filter(team %in% r2_winners) -> four_factors
four_factors |>
  mutate(
    matchup = case_when(
      seed %in% c(1,16,8,9,5,12,4,13) ~ "1/16/8/9 vs 5/12/4/13",
      seed %in% c(6,11,3,14,7,10,2,15) ~ "6/11/3/14 vs 7/10/2/15",
    )
  ) -> four_factors

four_factors

ggplot(four_factors, aes(x = eFG_pct.x, y = FT_FGA.x, color = as.factor(seed), group = matchup)) + geom_point(aes(size = ORtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()

ggplot(four_factors, aes(x = TOV_pct.x, y = ORB_pct.x, color = as.factor(seed), group = matchup)) + geom_point(aes(size = ORtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()

ggplot(four_factors, aes(x = eFG_pct.y, y = FT_FGA.y, color = as.factor(seed), group = matchup)) + geom_point(aes(size = DRtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()

ggplot(four_factors, aes(x = TOV_pct.y, y = ORB_pct.y, color = as.factor(seed), group = matchup)) + geom_point(aes(size = DRtg, shape = higher_seed)) + facet_wrap(~region) + geom_line()
```

```{r}
r3_winners = c("Connecticut","Illinois","James Madison","Kentucky",
               "Saint Mary's (CA)","New Mexico","Purdue",
               "Tennessee")

four_factors |> filter(team %in% r3_winners) -> four_factors
four_factors |>
  mutate(
    matchup = case_when(
      seed %in% c(1,16,8,9,5,12,4,13) ~ "1/16/8/9 vs 5/12/4/13",
      seed %in% c(6,11,3,14,7,10,2,15) ~ "6/11/3/14 vs 7/10/2/15",
    )
  ) -> four_factors

four_factors

ggplot(four_factors, aes(x = eFG_pct.x, y = FT_FGA.x, color = as.factor(team))) + geom_point(aes(size = ORtg, shape = higher_seed)) + facet_wrap(~region)

ggplot(four_factors, aes(x = TOV_pct.x, y = ORB_pct.x, color = as.factor(team))) + geom_point(aes(size = ORtg, shape = higher_seed)) + facet_wrap(~region)

ggplot(four_factors, aes(x = eFG_pct.y, y = FT_FGA.y, color = as.factor(team))) + geom_point(aes(size = DRtg, shape = higher_seed)) + facet_wrap(~region)

ggplot(four_factors, aes(x = TOV_pct.y, y = ORB_pct.y, color = as.factor(team))) + geom_point(aes(size = DRtg, shape = higher_seed)) + facet_wrap(~region)
```
```{r}
r4_winners = c("Connecticut","Kentucky",
               "Saint Mary's (CA)",
               "Tennessee")

four_factors |> filter(team %in% r4_winners) -> four_factors
four_factors |>
  mutate(
    matchup = case_when(
      seed %in% c(1,16,8,9,5,12,4,13) ~ "1/16/8/9 vs 5/12/4/13",
      seed %in% c(6,11,3,14,7,10,2,15) ~ "6/11/3/14 vs 7/10/2/15",
    )
  ) -> four_factors

four_factors

ggplot(four_factors, aes(x = eFG_pct.x, y = FT_FGA.x, color = as.factor(team))) + geom_point(aes(size = ORtg, shape = higher_seed))

ggplot(four_factors, aes(x = TOV_pct.x, y = ORB_pct.x, color = as.factor(team))) + geom_point(aes(size = ORtg, shape = higher_seed))

ggplot(four_factors, aes(x = eFG_pct.y, y = FT_FGA.y, color = as.factor(team))) + geom_point(aes(size = DRtg, shape = higher_seed))

ggplot(four_factors, aes(x = TOV_pct.y, y = ORB_pct.y, color = as.factor(team))) + geom_point(aes(size = DRtg, shape = higher_seed))
```

```{r}
kenpom <- read_csv("/Users/joshlafronz/Documents/my stuff/projects/march/kenpom.csv")
unique(kenpom$Conf)
```


```{r}
kenpom[grepl("[[:digit:]]", kenpom$Team),] |>
  separate(Team, c("team", "seed"), sep = "[0-9]+")
```


```{r}
str_trim(strsplit((kenpom[grepl("[[:digit:]]", kenpom$Team),])$Team, split = "[0-9]+"))
```

```{r}
kenpom[grepl("[[:digit:]]", kenpom$Team),] |>
  mutate(
    Team = str_trim(strsplit(Team,split = "[0-9]+"))
  ) -> kenpom
```

```{r}
kenpom$Team[kenpom$Team == "Iowa St."] <- "Iowa State"
kenpom$Team[kenpom$Team == "Michigan St."] <- "Michigan State"
kenpom$Team[kenpom$Team == "BYU"] <- "Brigham Young"
kenpom$Team[kenpom$Team == "Saint Mary's"] <- "Saint Mary's (CA)"
kenpom$Team[kenpom$Team == "San Diego St."] <- "San Diego State"
kenpom$Team[kenpom$Team == "Mississippi St."] <- "Mississippi State"
kenpom$Team[kenpom$Team == "Colorado St."] <- "Colorado State"
kenpom$Team[kenpom$Team == "Boise St."] <- "Boise State"
kenpom$Team[kenpom$Team == "Washington St."] <- "Washington State"
kenpom$Team[kenpom$Team == "Utah St."] <- "Utah State"
kenpom$Team[kenpom$Team == "McNeese St."] <- "McNeese State"
kenpom$Team[kenpom$Team == "Morehead St."] <- "Morehead State"
kenpom$Team[kenpom$Team == "South Dakota St."] <- "South Dakota State"
kenpom$Team[kenpom$Team == "Long Beach St."] <- "Long Beach State"
kenpom$Team[kenpom$Team == "Montana St."] <- "Montana State"
kenpom$Team[kenpom$Team == "Grambling St."] <- "Grambling"
kenpom$Team[kenpom$Team == "Charleston"] <- "College of Charleston"
kenpom$Team[kenpom$Team == "N.C. State"] <- "NC State"
kenpom
```

```{r}
kenpom |> filter(Team %in% field_of_68$team)
```

```{r}
field_of_68 |> anti_join(kenpom, by = join_by("team" == "Team"))
```

```{r}
kenpom |> relocate(Rk, .before = AdjO) -> kenpom
```

```{r}
kenpom |> left_join(advanced_off |> left_join(advanced_def, by = (c('seed', 'team', 'conference'))) |> subset(select=c(seed, team, conference, ORtg, DRtg))  |> mutate(
  net_rtg = ORtg - DRtg
), by = join_by("Team" == "team")) |> subset(select = c(Team, conference, AdjEM, net_rtg)) |>
  mutate(
    power_5 = case_when(
      conference %in% c("Big East", "Big 12", "Big Ten", "SEC", "ACC", "Pac-12") ~ TRUE,
      TRUE ~ FALSE
    )
  ) |> ggplot(aes(x = net_rtg, y = AdjEM, color = power_5)) + geom_point()
```

```{r}
true_shooting <- function(fga, fta, pts){
  tsa = fga + 0.44*fta
  ts_pct = pts/(2*tsa)
  print(ts_pct)
}

true_shooting(54,6,69)
```

```{r}
basic_off = read_csv("/Users/joshlafronz/Documents/my stuff/projects/datasets/march 23/basic_off.csv")
basic_def = read_csv("/Users/joshlafronz/Documents/my stuff/projects/datasets/march 23/basic_def.csv")
advanced_off = read_csv("/Users/joshlafronz/Documents/my stuff/projects/datasets/march 23/advanced_off.csv")
advanced_def = read_csv("/Users/joshlafronz/Documents/my stuff/projects/datasets/march 23/advanced_def.csv")
```

# Working with data from last year

```{r}
basic_off |>
  rename(tot_W = W...4, tot_L = L...5, conf_W = W...10, conf_L = L...11, home_W
         = W...13, home_L = L...14, away_W = W...16, away_L = L...17,
         PS = Tm., PA = Opp., team = School) |>
  subset(select = -c(Rk, ...9, ...12, ...15, ...18, ...21)) |>
  filter(grepl("NCAA", team) & 
           !team %in% c("Nevada NCAA","Texas Southern NCAA",
                        "Mississippi State NCAA", "Southeast Missouri State NCAA")) |>
  mutate(
    team = substr(team,1,nchar(team) - 5)
  ) -> basic_off

basic_def |>
  rename(tot_W = W...4, tot_L = L...5, conf_W = W...10, conf_L = L...11, home_W
         = W...13, home_L = L...14, away_W = W...16, away_L = L...17,
         PS = Tm., PA = Opp., team = School) |>
  subset(select = -c(Rk, ...9, ...12, ...15, ...18, ...21)) |>
  filter(grepl("NCAA", team) & 
           !team %in% c("Nevada NCAA","Texas Southern NCAA",
                        "Mississippi State NCAA", "Southeast Missouri State NCAA")) |>
  mutate(
    team = substr(team,1,nchar(team) - 5)
  ) |>
  rename(opp_FG = FG, opp_FGA = FGA, opp_FG_pct = FG_pct, opp_X3P = X3P, 
         opp_X3PA = X3PA, opp_X3P_pct = X3P_pct, opp_FT = FT, opp_FTA = FTA,
         opp_FT_pct = FT_pct, opp_ORB = ORB, opp_TRB = TRB, opp_AST = AST, opp_STL = STL,
         opp_BLK = BLK, opp_TOV = TOV, opp_PF = PF) -> basic_def

advanced_off |>
  rename(tot_W = W...4, tot_L = L...5, conf_W = W...10, conf_L = L...11, home_W
         = W...13, home_L = L...14, away_W = W...16, away_L = L...17,
         PS = Tm., PA = Opp., team = School) |>
  subset(select = -c(Rk, ...9, ...12, ...15, ...18, ...21)) |>
  filter(grepl("NCAA", team) & 
           !team %in% c("Nevada NCAA","Texas Southern NCAA",
                        "Mississippi State NCAA", "Southeast Missouri State NCAA")) |>
  mutate(
    team = substr(team,1,nchar(team) - 5)
  ) -> advanced_off

advanced_def |>
  rename(tot_W = W...4, tot_L = L...5, conf_W = W...10, conf_L = L...11, home_W
         = W...13, home_L = L...14, away_W = W...16, away_L = L...17,
         PS = Tm., PA = Opp., team = School, DRtg = ORtg) |>
  subset(select = -c(Rk, ...9, ...12, ...15, ...18, ...21)) |>
  filter(grepl("NCAA", team) & 
           !team %in% c("Nevada NCAA","Texas Southern NCAA",
                        "Mississippi State NCAA", "Southeast Missouri State NCAA")) |>
  mutate(
    team = substr(team,1,nchar(team) - 5)
  ) |>
  rename(opp_FTr = FTr, opp_X3PAr = X3PAr, opp_TS_pct = TS_pct, opp_TRB_pct = TRB_pct,
         opp_AST_pct = AST_pct, opp_STL_pct = STL_pct, opp_BLK_pct = BLK_pct,
         opp_eFG_pct = eFG_pct, opp_TOV_pct = TOV_pct, opp_ORB_pct = ORB_pct, opp_FT_FGA =
           FT_FGA) -> advanced_def
```

```{r}
# basic_off
# basic_def
basic_off |> left_join(basic_def, 
                       by = c("team","G","tot_W","tot_L","W_L_pct", "SRS","SOS","conf_W","conf_L","home_W", "home_L","away_W","away_L","PS","PA","MP")) -> basic_off_def
basic_off_def
```

```{r}
# advanced_off
# advanced_def
advanced_off |> left_join(advanced_def, by = c("team","G","tot_W","tot_L","W_L_pct", "SRS","SOS","conf_W","conf_L","home_W", "home_L","away_W","away_L","PS","PA","Pace")) -> advanced_off_def
advanced_off_def
```

```{r}
# basic_off_def
# advanced_off_def
basic_off_def |> left_join(advanced_off_def, by = c("team","G","tot_W","tot_L","W_L_pct", "SRS","SOS","conf_W","conf_L","home_W", "home_L","away_W","away_L","PS","PA")) -> final_table
rm(basic_off, basic_def, advanced_off, advanced_def, basic_off_def, advanced_off_def)
final_table
```

```{r}
r_32 = c("FDU","Florida Atlantic","Duke","Tennessee","Kentucky","Kansas State",
         "Michigan State","Marquette","Houston","Auburn","Miami (FL)","Indiana",
         "Pittsburgh","Xavier","Penn State","Texas","Alabama","Maryland","San Diego State",
         "Furman","Creighton","Baylor","Missouri","Princeton","Kansas","Arkansas",
         "Saint Mary's (CA)","Connecticut","TCU","Gonzaga","Northwestern","UCLA")

s_16 = c("Florida Atlantic","Tennessee","Kansas State","Michigan State", "Houston",
         "Miami (FL)","Xavier","Texas","Alabama","San Diego State","Creighton",
         "Princeton","Arkansas","Connecticut","Gonzaga","UCLA")

e_8 = c("Florida Atlantic","Kansas State","Miami (FL)","Texas","San Diego State",
        "Creighton","Connecticut","Gonzaga")

final_4 = c("Florida Atlantic","Miami (FL)", "San Diego State", "Connecticut")

final = c("San Diego State","Connecticut")

# winner = "Connecticut"

final_table |> filter(team %in% r_32) |> filter(team %in% s_16) |> 
  filter(team %in% e_8) |> filter(team %in% final_4) |> filter(team %in% final) |>
  filter(team == "Connecticut")
```

```{r}
final_table |> mutate(
  r_32 = ifelse(team %in% r_32, 1, 0),
  s_16 = ifelse(team %in% s_16, 1, 0),
  e_8 = ifelse(team %in% e_8, 1, 0),
  final_4 = ifelse(team %in% final_4, 1, 0),
  final = ifelse(team %in% final, 1, 0),
  winner = ifelse(team == "Connecticut", 1, 0),
) -> final_table
rm(r_32, s_16, e_8, final_4, final)
final_table
```
```{r}
final_table
```

```{r}
#specify the full model with all the potential predictors
#and the null model with the intercept only

## trying with every single variable

# res.full=lm(ORtg ~ AGE + GP + MPG + USG. + TO. + FTA + FT. + X2PA + X2P. + 
#                 X3PA + X3P. + eFG. + TS. + PPG + RPG + APG + SPG + BPG + TPG + 
#                 P.R + P.A + P.R.A + VI + DRtg, data=nba.data)
# 
# res.null=lm(ORtg~1,data=nba.data)

res.full=glm(final ~ G + tot_W + tot_L + W_L_pct + SRS + SOS + conf_W + conf_L +
               home_W + home_L + away_W + away_L + PS + PA + MP + FG + FGA + FG_pct +
               X3P + X3PA + X3P_pct + FT + FTA + FT_pct + ORB + TRB + AST + STL + 
               BLK + TOV + PF + opp_FG + opp_FGA + opp_FG_pct + opp_X3P + opp_X3PA + 
               opp_X3P_pct + opp_FT + opp_FTA + opp_FT_pct + opp_ORB + opp_TRB +
               opp_AST + opp_STL + opp_BLK + opp_TOV + opp_PF + Pace + ORtg + FTr +
               X3PAr + TS_pct + TRB_pct + AST_pct + STL_pct + BLK_pct + eFG_pct + 
               TOV_pct + ORB_pct + FT_FGA + DRtg + opp_FTr + opp_X3PAr + opp_TS_pct + 
               opp_TRB_pct + opp_AST_pct + opp_STL_pct + opp_BLK_pct + opp_eFG_pct + 
               opp_TOV_pct + opp_ORB_pct + opp_FT_FGA, data = final_table)

res.null = glm(final ~ 1, data = final_table)
```

```{r}
res.full=glm(final ~ W_L_pct + SRS + SOS + conf_W + conf_L + home_W + home_L + away_W + away_L + FG_pct + X3P_pct + FT_pct + opp_FG_pct + opp_X3P_pct + opp_FT_pct + Pace + ORtg + FTr + X3PAr + TS_pct + TRB_pct + AST_pct + STL_pct + BLK_pct + eFG_pct + TOV_pct + ORB_pct + FT_FGA + DRtg + opp_FTr + opp_X3PAr + opp_TS_pct + opp_TRB_pct + opp_AST_pct + opp_STL_pct + opp_BLK_pct + opp_eFG_pct + opp_TOV_pct + opp_ORB_pct + opp_FT_FGA, data = final_table)

res.null = glm(final ~ 1, data = final_table)
```

```{r}
# forward=step(res.null,
#              scope = list(upper=res.full),
#              direction="forward",
#              test="F",
#              data=final_table)

backward=step(res.full,
              scope = list(lower=res.null),
              direction="backward",
              test="F",
              data=final_table)
```

```{r}
forward=step(res.null,
             scope = list(upper=res.full),
             direction="forward",
             test="F",
             data=final_table)
```


```{r}
# forward=step(res.null,
#              scope = list(upper=res.full),
#              direction="forward",
#              test="F",
#              data=final_table)

# res = glm(final ~ G + opp_X3P + opp_FT_pct + opp_STL_pct + opp_FGA + TOV_pct + 
#     STL_pct + BLK + opp_STL + X3P + opp_FTr + opp_FT + opp_FTA + 
#     opp_FT_FGA + DRtg, data=final_table)

res = glm(final ~ W_L_pct + SRS + SOS + conf_W + conf_L + home_L + opp_X3P_pct + 
    opp_FT_pct + ORtg + FTr + TS_pct + STL_pct + BLK_pct + eFG_pct + 
    FT_FGA + DRtg + opp_TS_pct + opp_AST_pct + opp_STL_pct + 
    opp_eFG_pct + opp_TOV_pct + opp_ORB_pct + opp_FT_FGA, data = final_table)

summary(res)

extractAIC(res)

ols_mallows_cp(res,res.full)

res = glm(final ~ opp_X3P_pct + opp_FT_pct, data = final_table)

summary(res)

extractAIC(res)

ols_mallows_cp(res,res.full)
```
```{r}
glm_eval <- function(model, full) {
  print(summary(model))
  print(extractAIC(model))
  print(vif(model))
  print(ols_mallows_cp(model, full))
}

res = glm(final ~ W_L_pct + SRS + SOS + conf_W + conf_L + home_L + opp_X3P_pct +
    opp_FT_pct + ORtg + FTr + TS_pct + STL_pct + BLK_pct + eFG_pct +
    FT_FGA + DRtg + opp_TS_pct + opp_AST_pct + opp_STL_pct +
    opp_eFG_pct + opp_TOV_pct + opp_ORB_pct + opp_FT_FGA, data = final_table)

glm_eval(res, res.full)

res = glm(final ~ W_L_pct + SRS + SOS + conf_W + conf_L + home_L + opp_X3P_pct +
    opp_FT_pct + ORtg + FTr + TS_pct + STL_pct + BLK_pct + eFG_pct +
    FT_FGA + DRtg + opp_TS_pct + opp_STL_pct +
    opp_eFG_pct + opp_TOV_pct + opp_ORB_pct + opp_FT_FGA, data = final_table)

glm_eval(res, res.full)

res = glm(final ~ W_L_pct + SRS + SOS + conf_W + conf_L + home_L + opp_X3P_pct +
    opp_FT_pct + ORtg + TS_pct + STL_pct + BLK_pct + eFG_pct +
    FT_FGA + DRtg + opp_TS_pct + opp_STL_pct +
    opp_eFG_pct + opp_TOV_pct + opp_ORB_pct + opp_FT_FGA, data = final_table)

glm_eval(res, res.full)

res = glm(final ~ W_L_pct + SRS + SOS + conf_W + conf_L + home_L + opp_X3P_pct +
    opp_FT_pct + ORtg + TS_pct + BLK_pct + eFG_pct +
    FT_FGA + DRtg + opp_TS_pct + opp_STL_pct +
    opp_eFG_pct + opp_TOV_pct + opp_ORB_pct + opp_FT_FGA, data = final_table)

glm_eval(res, res.full)


```

```{r}
#start with all potential predictors in the model

# backward=step(res.full,
#               scope = list(lower=res.null),
#               direction="backward",
#               test="F",
#               data=final_table)

# res1 = lm(ORtg ~ MPG + USG. + TO. + FTA + FT. + X3P. + eFG. + TS. + PPG + 
#             RPG + APG + SPG + TPG + P.R, data=nba.data)
# 
# summary(res1)
# 
# extractAIC(res1)
# 
# ols_mallows_cp(res1,res.full)
```










